\documentclass[a4paper,12pt]{report}

\usepackage{enumitem}  % for the leftmargin option
\usepackage{etaremune}  % reverse enumerate for papers, etc.
\usepackage[
    colorlinks=true, 
    linkcolor=black, % TOC
    urlcolor=blue, 
    citecolor=blue
]{hyperref}  %  for hyperlinks

% putting in citations via a bib file with bold names
\usepackage[
    backend=biber,
    style=numeric,
	sorting=none,
	maxbibnames=99,
	maxcitenames=99,
	abbreviate=false
  ]{biblatex}
  
\DeclareNameAlias{author}{last-first}
\addbibresource{NLaws.bib}
% for excluding specific entries from printbibliography:
\DeclareBibliographyCategory{fullcited}
\newcommand{\mybibexclude}[1]{\addtocategory{fullcited}{#1}}
\DeclareFieldFormat{title}{\iffieldundef{url}{#1}{\href{\thefield{url}}{\color{blue}#1}}}


\title{\Huge \textbf{Curriculum Vitae\\Nicholas D. Laws, Ph.D.} }
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
\maketitle

\tableofcontents
\newpage

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Professional Experience}
 % Add to ToC with hyperlink
\addcontentsline{toc}{section}{\protect\numberline{}Professional Experience} 
\begin{itemize}[leftmargin=*]
    \item Director of Modeling and Analytics, IdeaSmiths LLC, 2024--present
    \item Senior Power Systems Optimization Engineer, Camus Energy, 2022--present
    \item Research Engineer, National Renewable Energy Laboratory, 2017--2022
    \item Graduate Research Assistant, Webber Energy Group, UT Austin, 2018--2023
    \item Graduate Research Assistant, Thayer School of Engineering, Dartmouth College, 2014-2016
    \item Mountain, Climbing, and Ski Guide, Aspen Alpine Guides, 2013-2014
    \item Mountain Guide, International Mountain Guides, 2011-2014
    \item Commodity Buyer, Valley Lumber, 2007-2011
    \item Construction Manager, Eldorado Climbing Walls, 2006-2007
\end{itemize}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Education}
\addcontentsline{toc}{section}{\protect\numberline{}Education}
\begin{itemize}[leftmargin=*]
    \item Ph.D. Mechanical Engineering, The University of Texas at Austin, 2023
    \item M.S. Engineering Sciences, Thayer School of Engineering Dartmouth College, 2016
    \item B.S. Aerospace Engineering, Boston University, 2005
\end{itemize}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Peer-Reviewed Journal Articles}
\addcontentsline{toc}{section}{\protect\numberline{}Peer-Reviewed Journal Articles}
\begin{etaremune}
    \item \fullcite{laws2023der4nwa}\mybibexclude{laws2023der4nwa}
    \item \fullcite{laws2022linearizing}\mybibexclude{laws2022linearizing}
	\item \fullcite{mishra2021computational}\mybibexclude{mishra2021computational}
	\item \fullcite{mclaren2019solar}\mybibexclude{mclaren2019solar}
	\item \fullcite{laws2018impacts}\mybibexclude{laws2018impacts}
	\item \fullcite{anderson2018quantify}\mybibexclude{anderson2018quantify}
	\item \fullcite{laws2017utility}\mybibexclude{laws2017utility}
	\item \fullcite{laws2016hydrokinetic}\mybibexclude{laws2016hydrokinetic}
\end{etaremune}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Peer-Reviewed Conference Proceedings}
\addcontentsline{toc}{section}{\protect\numberline{}Peer-Reviewed Conference Proceedings}
\begin{etaremune}
    \item \fullcite{gasper2023conf}\mybibexclude{gasper2023conf}
    \item \fullcite{mirletz2023impacts}\mybibexclude{mirletz2023impacts}
    \item \fullcite{laws2017singularity}\mybibexclude{laws2017singularity}
\end{etaremune}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Non-Refereed Conference Papers and Presentations}
\addcontentsline{toc}{section}{\protect\numberline{}Non-Refereed Conference Papers and Presentations}
\begin{etaremune}
    \item \fullcite{laws2018naesco}\mybibexclude{laws2018naesco}
\end{etaremune}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Selected Technical Reports and White Papers}
\addcontentsline{toc}{section}{\protect\numberline{}Select Technical Reports and White Papers}
\begin{etaremune}
    \item \fullcite{laws2023stoch}\mybibexclude{laws2023stoch}
    \item \fullcite{murphy2020adapting}\mybibexclude{murphy2020adapting}
    \item \fullcite{mclaren2018valuing}\mybibexclude{mclaren2018valuing}
\end{etaremune}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Selected Projects and Roles}
\addcontentsline{toc}{section}{\protect\numberline{}Selected Projects}


\textbf{Director of Modeling and Analytics -- IdeaSmiths $|$ 2024--present}

IdeaSmiths delivers actionable insight for its clients. From deep dives on technical due diligence
to high-level assessments of market trends, IdeaSmiths helps its clients understand current and future
challenges, and opportunities for the energy industry. As Director of Modeling and Analytics, 
Dr. Laws keeps IdeaSmiths at the forefront of industry tools and methods for:
\begin{itemize}
    \item Energy Technology Analysis and Due Diligence
    \item Integrated Systems Modeling and Analysis
    \item Expert Witness Analysis and Testimony
    \item Prototype Development and Testing
\end{itemize}


\noindent
\textbf{Senior Power Systems Optimization Engineer -- Camus Energy $|$ 2022--present}

At Camus Energy Dr. Laws develops novel software to aid electric utilities in the energy transition,
including:
\begin{itemize}
    \item Automation of utility data into power flow models, including SCADA, AMI, and GIS.
    \item Novel graph methods to determine system connectivity of power distribution systems.
    \item Optimal power flow methods for dynamic operating envelopes to support flexible
    interconnection of new storage and generation assets.
    \item Optimal inverter based resource set points to support utility objectives
\end{itemize}

\noindent
\textbf{Textbook: Practical Optimal Power Flow $|$ 2023-present}

Building on his novel theoretical contributions from his Ph.D. research and extensive practical
experience, Dr. Laws is writing a text book on Practical Optimal Power Flow. The book is the first
to extensively cover multiphase methods for modeling power distribution networks, as well as
techniques for modeling inverter based resources. Methods for optimization under uncertainty are also
covered -- including Decision Making Under Uncertainty techniques for developing solutions that are
robust against unknown unknowns.
\newline
\newline
\noindent
\textbf{Battery State-of-Health informed optimal dispatch and sizing $|$ 2022}

Using an empirical battery degradation model Dr. Laws incorporated depth of discharge and equivalent
full cycle decision variables into a mixed integer linear program to account for battery
state-of-health. The state-of-health variable allowed for the incorporation of battery maintenance
costs as well as accounting for the impacts of hourly dispatch decisions on battery degradation.
\newline
\newline
\noindent
\textbf{REopt Lead Developer -- NREL $|$ 2021-2022}

Dr. Laws led the transition of REopt to an open-source Julia package and Python-based API to support
a web tool (\href{http://reopt.nrel.gov/tool}{reopt.nrel.gov/tool}) as well as make the REopt tool
more accessible and customizable.
\newline
\newline
\noindent
\textbf{Co-optimizing distributed energy resources for system operator and customer objectives $|$ 2020}

Dr. Laws acted as Principal Investigator on a competitively awarded internal research and
development grant at NREL on novel methods to determine optimal capacities and economic incentives
for distributed energy resources from multiple perspectives.
\newline
\newline
\noindent
\textbf{Stochastic Price Generation for Evaluating Wholesale Electricity Mar-
ket Bidding Strategies $|$ 2019}

Dr. Laws led the development of a novel technique for generation wholesale electricity market price
scenarios to support stochastic optimization models that determine optimal bidding strategies for
energy storage assets.
\newline
\newline
\noindent
\textbf{Value of Resiliency Laboratory Directed R\&D $|$ 2018}

Dr. Laws supported an NREL internally funded research and development project to determine methods for
valuing the resiliency of energy storage and generation assets for all stakeholders, including asset
owners, grid operators, and society.
\newline
\newline
\noindent
\textbf{National Scale Analysis of the Economics of Commercial Behind-The-Meter Solar Plus Storage $|$ 2017}

Dr. Laws' first project at NREL required adapting and scaling an in-house analysis tool that
required an expert analyst weeks to months to curate many inputs and analyze the outputs into a tool
that could be automated for tens-of-thousands of runs to support a national scale analysis of where
in the United States behind-the-meter solar plus storage was economical for commercial building
owners.


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Skills}
\addcontentsline{toc}{section}{\protect\numberline{}Skills}
\begin{itemize}
    \item Expert in Mathematical Programming including Optimization Under Uncertainty, Integer Programming, Optimal Control, Optimal Power Flow, and Bilevel optimization
    \item Familiar with many energy and ancillary service markets, power system regulatory bodies, and the deregulation of electric utilities.
    \item Familiar with a wide range power flow formulations, including relaxations and approximations for handling large optimization problems.
    \item Expert in Python, Julia, Docker, Git, test-driven development
    \item Experience with commercial and free solvers including Gurobi, CPLEX, Press, HiGHS, and Ipopt.
    \item Experience with many mathematical programming languages including JuMP, CVXPY, and Pyomo.
    \item Experience with full stack development, including Javascript, HTML, RESTful API, Kubernetes, PostgreSQL, Jenkins, multi-threaded task queues, AWS, and GCP.
    \item Contributor and author of several open source optimization packages ({\href{http://github.com/nlaws}{github.com/nlaws}}).
    \item Experience with OpenDSS for steady state power flow analysis including hosting capacity studies. 
\end{itemize}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Biography}
\addcontentsline{toc}{section}{\protect\numberline{}Biography}
Dr. Nick Laws is the Director of Modeling and Analytics at IdeaSmiths. He is also the Senior Power
Systems Optimization Engineer at Camus Energy, where he researches and develops methods and software
to aid electric utilities in the energy transition. Prior to Camus Energy, Dr. Laws led the
development of a flagship decision support tool at the National Renewable Energy Laboratory (NREL),
called REopt. REopt is a techno-economic optimization tool for determining distributed energy
resource (DER) capacities and dispatch strategies to meet economic, resilience, and emission
reduction goals.  Dr. Laws is the sole-author of the 2022 “Centralized Control of DER with Price
Signals” NREL software record, developed to control diverse DER with minimal information exchange.
Dr. Laws also served as Principal Investigator on the 2020 NREL Seed Laboratory Directed R\&D project
“Co-optimizing DER for system operator and customer objectives”. Dr. Laws is passionate about
preserving wild spaces for future generations and enjoying wilderness with family and friends.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \printbibliography[notcategory=fullcited]
\end{document}